Stack Safety as a Security Property

### 2020-07-09 main.tex

\paragraph*{Capability machines}
%
Other policies are implemented in hardware or built on such hardware mechanisms.
Capability machines like CHERI \rb{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies involving, for
example, compartmentalization or the stack protection policies of
\citet{Skorstengaard+19,Skorstengaard+19b}. Mapping the underlying protection
mechanisms to policy states and step function would be a delicate operation, as
those are tightly coupled to the hardware design. However, separating the
enforcement aspects of a concrete policy (such as StkTokens, where \(\POLS\)
would encompass details like the abstract call stack, the capability sealing
scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
the machine would be relatively straightforward.

@@

%Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
%We then define {\em policy-sensitive observational equivalence}: a
%trace $\MPT \in \MPOTRACE$ is equivalent to a trace $\machT \in
%\MOTRACE$ if (1) \(\MPT\) is infinite and the observations of \(\MPT\)
%and \(\machT\) are equivalent, or (2) if the \(\MPT\) is finite (due to a
%policy fault) and its observations are a prefix of those of M.
%
%% Fixed by Sean
%%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
%\judgmenttwo{\(\neg \isinf(\MPT)\)}{\(\pi_o(\MPT) \lesssim \pi_o(\machT)\)}
%            {\(\MPT \sim_P \machT\)}
%
%\judgmenttwo{\(\isinf(\MPT)\)}{\(\pi_o(\MPT) \simeq \pi_o(\machT)\)}
%            {\(\MPT \sim_P \machT\)}

@@

When \(\mach_c\) is a call state (\(\callmap(\mach_c(\PCname)) = n\) for some $n$) and
\(\mach_c \stepstoobs{\obs} \mach_e\) for some 
$\mach_e$ and $\obs$, we can construct a contour based on the bounds of the
callee's stack frame.
%
The private state of the callee is the entire area of the memory from
stack base to stack pointer, with the exception of the $n$ locations
corresponding to the arguments and result locations of the
call. Private components are marked high confidentiality and high
integrity: the callee should be able to neither read nor write them.
%
Everything above the stack pointer at the point of the call is
``uninitialized'' memory that can be written into but should not be
read. To protect the callee itself from such reads, these components
are marked high confidentiality and low integrity.
%
The same is true for the designated output location: it can be
freely written into but should not be read.
%
Special care is needed for the code memory corresponding to the
callee: they are marked low confidentiality if they belong to the
callee and high confidentiality otherwise; regardless they are always
high integrity.\apt{Sean:I think this is vestigial from an aborted version
  in which you were modeling WBCF using confidentiality. In any casen
  it is not reflected in the definition below.}
%
Finally, registers and the frame of arguments and results are free
for the callee to use, and are therefore tagged with low integrity and
low confidentiality.

@@

\definition
Let $\contour$ be a contour and $\MPT$ be a (potentially infinite)
$\MPOTRACE$. We will say $\MPT$ satisfies {\em trace integrity} with
respect to $\contour$ if its final state (if any) agrees with its
initial one on all components marked $\HIGHSEC$ in $\contour$:
%
\[\begin{split}
  \mathit{int} ~ \contour ~ \MPT \triangleq &
    \head(\MPT) = \mpostate{\mach_e}{\_}{\_}  \land 
    \last(\MPT) = \mpostate{\mach_r}{\_}{\_} \Rightarrow \\
  & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
    \mach_e(\component) = \mach_r(\component) \\
\end{split}\]
%\leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
%\sna{We could do it like this, with finiteness implicit in the existence of
%     a final state.}+1

@@

Trace integrity characterizes the trace corresponding to a single
callee.
Going back to our running example, such a trace would be, for
instance, the entire trace of {\tt h} from instruction 40 to
instruction 47. 
As an example of an integrity violation affecting this trace, \cref{fig:intex} shows what happens
if we change line 44 of the example program to $[\rsp - 5] \leftarrow r_4$.
Instead of storing data to the return value slot,
the changed version overwrites {\tt g}'s local variable {\tt z}.
The figure shows the modified trace, with snapshots of the initial
and final states colored to show the contour; high 
integrity memory is in red\apt{NO?}, while low integrity is in green\apt{NO?}.
At the end of the trace high integrity address 103 should match the initial state, and does not.
So trace integrity is violated.
\rb{There are some color name macros, but none for integrity yet? Use instead
  of hardcoded color names.}\apt{I don't understand what's going on with the colors.
  I thought we agreed to use the same coloring scheme throughout?}

We now lift this definition to an integrity property for a complete
execution trace. We do this by requiring that trace integrity holds
for every subtrace delimited by a call and its corresponding return.
%for every transition between a caller and a callee for the entirety of
%the trace of the callee.\apt{?????}

\apt{Some more words needed here to smooth the technicalities.}

\definition%{Transition}
Given a trace $\MPT$, an $n$-{\em transition} is a pair of
machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
in the trace such that:
\begin{itemize}
\item one state pair steps to the other:
  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstopol{\obs} (\mach_e, \pol_e),$$
\item $\mach_c$ is a valid call with $n$ arguments:
  $$\callmap(\mach_c(\PCname)) = n,$$
\item and $\mach_e$ is an entry point\apt{or is in some function?? and why do we need either??}:
  $$\exists f. \codemap(\mach_e(\PCname)) = f.$$
\end{itemize}

\definition%{Stack Integrity}
A system satisfies {\em stack
  integrity} if, for every
initial machine-policy state pair $(\mach_0, \pol_0)$ and its induced
trace $\MPT$, and every $n$-transition $(\mach_c, \pol_c)$ and
$(\mach_e, \pol_e)$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from
$(\mach_e, \pol_e)$ until the first return corresponding to $\mach_c$
(if any) satisfies trace integrity with respect to the contour at the
call:
%
$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
\MPT_{pre} ~ | ~ \ret{\mach_c} \Rightarrow
\mathit{int}~\mathit{Cof}(\mach_c, n)~\MPT_{pre}$$

@@

Unlike integrity, confidentiality is a noninterference-style property,
comparing the behavior of a machine with an arbitrary variant that
preserves low confidentiality data.

\definition%{Variations}
We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
another with respect to a contour $\contour$ if they agree on low
confidentiality data:
%
\[
\mach \approx_\contour \mach' \triangleq \forall \component .
\contour(\component) = \LOWSEC \Rightarrow \mach(\component) =
\mach'(\component)
\]
%
Then, the trace version of confidentiality operates on two traces: the
``original'' tracce $MP \in \MPOTRACE$ that corresponds to a callee $f$
and a ``variant'' trace $M \in \MTRACE$ that corresponds to how $f$
would behaves if it were started from a variant initial machine state.
%
As an example of variations, consider the top of \Cref{fig:confex1}
which shows a snapshot of the state as in the running example (top
memory layout) and a variation of it (bottom). \leo{Finalize
  contents, describe}

\definition%{Trace Confidentiality}
%% Suppose $m_e \approx_\contour m_e'$.
%% Two traces $MP$ with $\head(MP) = m_e$ and $M$ with $head(M) = m_e'$
%% satisfy {\em trace confidentiality} with respect to $\contour$ and
%% a return predicate \(R \subset \MACHS\) if:
Two traces $MP$, with head $m_e$, and $M$, with head $m_e'$,
satisfy {\em trace confidentiality} with respect to contour $\contour$ and
a return predicate \(R \subset \MACHS\) if, whenever $m_e \approx_\contour m_e'$ then:
\begin{itemize}
\item If $\MPT$ is finite with final state $\mach_r$ and $\mach_r \in R$, 
  then $\machT$ is also finite with some final state $\mach_r'$ and
  any changes from the initial states have changed in the same way:
$$\forall \component. ~ \mach_e(\component) \not = \mach_r(\component)
  \vee \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
  \mach_r(\component) = \mach_r'(\component)$$
  In addition their observations must be similar:
  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
\item If $\MPT$ is infinite, then $\machT$ is also infinite and their
  observations are similar:
  $$\pi_o(\MPT) \eqsim \pi_o(\machT)$$
\item If the $\MPT$ ends prematurely, without fulfilling \(R\), then
  we require that its observations are a prefix of those of $\machT$:
  $$\pi_o(\MPT) \lesssim \pi_o(\machT)$$
\end{itemize}
\apt{Needs words about why we need a return predicate abstraction.}

Consider the result of modifying line 42 of our running example to $r_4 \leftarrow [\rsp - 9]$.
Instead of reading the proper argument {\tt v}, {\tt h} now reads directly from the
secret value of {\tt x} that belongs to {\tt main}. \Cref{fig:confex}
shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
in {\high} and low confidentiality in \low. Below it, the variant state
\(\mach_e'\) agrees on low confidentiality memory, but all other addresses
hold -5, in orange.\apt{NO?}  After the last step, the return states \(\mach_r\) and
\(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
\(\mach_e'\). Addresses 108 and 109 have changed, shown in \tainted, and while 109
had changed in a consistent way (independent of the variation), 108 holds either 43 or -4, so the trace does not obey trace confidentiality.
\rb{Color name macros}
\apt{Explain slash notation in effects column.}

As another example, 
changing line 30 of the running example to $r_5  \leftarrow [\rsp - 6]$ would cause
{\tt g} to read the value of {\tt main}'s local variable {\tt x} and then immediately
output it, violating the internal observation clause of the confidentiality definition.
%\Cref{fig:confex2} shows the initial states \(\mach_e\), above,
%and \(\mach_e'\), below. Low confidentiality data in green is shared
%between them, while high confidentiality data in red and yellow may differ.
%After executing the code shown, the final states below have written different
%values to address 42, represented by the black and white components.
%It is allowed for address 40 to differ between the states due to the initial
%variation, but address 42 changed differently between the traces, representing
%a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
%is read, it is overwritten before return, and the write at step 3 always writes
%the same value. \Gprivc and red locations differ between variations only because
%of the initial variation, and the rest are identical. So this example passes.

@@

%\Cref{fig:confex2} gives an example of a trace that fails trace
%confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
%Since {\tt x}, at address 100, is high confidentiality (red) the variant
%may have a different value (yellow). So both \(r_4\) and address 107
%differ in the final state between the original trace and the variant one,
%failing the ending condition of trace confidentiality.

@@

\definition%{Stack Confidentiality}
%Let $\callmap$ be a call map and $\codemap$ a function map.
We say that a
system satisfies {\em stack confidentiality}
% with respect to $\callmap$ and $\codemap$
if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ 
in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)}
\mach_e\), let \(\machT\) be the similarly prefixed trace such that
\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, n)\)
and \(R = \ret{\mach_c}\):

$$\forall \MPT_{\mathit{call}} ~ \machT. ~
\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, n)} \mach_e' \wedge
\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \Rightarrow$$
$$\mathit{conf}~\mathit{Cof}(\mach_c, n) ~ (\ret{\mach_c}) ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}$$

\apt{oy. still needs work to make this intelligible}

@@

\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}
\newcommand*{\testInt}[3]{\ensuremath{\mathit{int}_T ~ #1 ~ #2 ~ #3}}
\newcommand*{\testConf}[2]{\ensuremath{\mathit{conf}_T ~ #1 ~ #2}}
\newcommand*{\testVS}[2]{\ensuremath{\mathit{\vs}_T ~ #1 ~ #2}}
\newcommand*{\testProp}[3]{\ensuremath{\mathit{stack}_T ~ #1 ~ #2 ~ #3}}

\definition
  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{stack integrity test} if their high-integrity components
  match:
  \[
    \testInt{\contour}{\mach}{\mach'} \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\component) = \mach'(\component)
  \]

\definition Two machine states $\mach_1$ and $\mach_2$ satisfy the
\emph{stack confidentiality test} if the machines they stepped to
have their changed components coincide and the observations during
these steps are equal:
\[
\begin{array}{rl}
  \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \triangleq &
  \exists \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2'. ~
  \mach_1 \stepstoobs{\obs_1} \mach_1' \wedge \mach_2 \stepstoobs{\obs_2} \mach_2' \Rightarrow \\
&    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
\end{array}
  \]

\definition The testing property keeps track of a variation machine
for every nested call performed so far on a {\em variant stack}: a
list of machine states annotated with a contour, each element of which
(denoted $\vm{C}$) of which is a variation of the actual machine
according to a contour $C$.
%
\footnote{The full formal definition in Coq also records additional
  information (the initial machine state at the call, its initial
  variation, the contour, and a return predicate) which we elide
  here for simplicity of exposition.}
%%
%  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
%  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
%  $$\vs \in \VSS$$
%%
\iflater
\leo{This can be elided}
  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
  respect to it if the stack is nonempty, the return map of the bottom element
  is the false property, and for each element $\vse \in \vs$:
  \begin{itemize}
  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
  \item $\mach$ is in the trace $\machT$ of its initial machine,
    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
  \item Its variant state $\mathit{curr\_variant}(\vse)$
    is in the trace of its initial variant,
    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item The registers of its contour $\contour(\vse)$ are public.
  \end{itemize}
  \fi

\definition
%
  Our testing property extends the basic tests above to whole variant
  stacks.  Given a machine that steps $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$, say that a variant stack $\vs$ \iflater that
  is well-formed with respect to $\mach$\fi safisfies the
  \emph{variant stack test} if for each variant element $\vm{C}$
  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step,
  denoted $\vm{C} \stepstoobs{\obs} \mach'$, while
  satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
  and stack confidentiality test $\testConf{\mach}{\mach'}$
  with regard to the original machine state. We write
  this $\testVS{(\mach, \pol)}{\vs}$.

\definition
%  Let $\callmap$ a call map and $\codemap$ a function map.
  Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
  variant stack $\vs$, we say that an MP-trace induced from $(\mach,
  \pol)$ satisfies the \emph{strong stack safety} at a given call
  depth as follows.
% Refactor repetitive notation before case split
  In all cases that let $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$.
  %and assume (unless the step is a return) that the
  %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
  %machines can step.

@@

  \item
%
    If the machine takes a \emph{call step} and $\contour$ is the contour at
    the call then,
%
    (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
    satisfied,
%
    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
    is satisfied with regard to an arbitrary $\contour$-variant
    of $\mach$,
%
    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
    $\contour$-variant of the machine,
%
    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
    \pol') : \MPT)}{vs}$.

>>>>

\paragraph*{Capability machines}
%
\rb{TODO Here the policy is built into the hardware, although the extended ISA
  could be split into a machine without capabilities and capabilities as the
  policy.}

@@

@@

When \(\mach_c\) is a call state (\(\callmap(\mach_c(\PCname)) = n\) for some $n$) and
\(\mach_c \stepstoobs{\obs} \mach_e\) for some 
$\mach_e$ and $\obs$, we can construct a contour based on the bounds of the
callee's stack frame.
%
The private state of the callee is the entire area of the memory from
stack base to stack pointer, with the exception of the $n$ locations
corresponding to the arguments and result locations of the
call. Private components are marked high confidentiality and high
integrity: the callee should be able to neither read nor write them.
%
Everything above the stack pointer at the point of the call is
``uninitialized'' memory that can be written into but should not be
read. To protect the callee itself from such reads, these components
are marked high confidentiality and low integrity.
%
The same is true for the designated output location: it can be
freely written into but should not be read.
%
Code is always high integrity and low confidentiality.
%
Finally, registers and the frame of arguments and results are free
for the callee to use, and are therefore tagged with low integrity and
low confidentiality.

@@

\definition

Let \(\contour\) be a contour and \(\MPT\) be a (potentially infinite)
\(\MPOTRACE\). We will say \(\MPT\) satisfies {\em trace integrity} with
respect to \(\contour\) if its final state (if any) agrees with its
initial one on all components marked \(\HIGHSEC\) in \(\contour\):
%
\[\begin{split}
  \propdef{TraceIntegrity} ~ \contour ~ \MPT \triangleq &
    \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
  & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
  & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
    \mach_e(\component) = \mach_r(\component) \\
\end{split}\]
\sna{Experimenting with font on the mathy version of the property.}

@@

Trace integrity characterizes the trace corresponding to a single
callee.
Going back to our running example, such a trace would be, for
instance, the entire trace of {\tt h} from instruction 40 to
instruction 47. 
As an example of an integrity violation affecting this trace, \cref{fig:intex} shows what happens
if we change line 44 of the example program to $[\rsp - 5] \leftarrow r_4$.
Instead of storing data to the return value slot,
the changed version overwrites {\tt g}'s local variable {\tt z}.
The figure shows the modified trace, with snapshots of the initial
and final states colored to show the contour; high 
integrity memory is in \high, while low integrity is in \low.
At the end of the trace high integrity address 103 should match the initial
state, and does not. So trace integrity is violated.
\rb{There are some color name macros, but none for integrity yet? Use instead
  of hardcoded color names.}\apt{I don't understand what's going on with the colors.
  I thought we agreed to use the same coloring scheme throughout?}

We now lift this definition to an integrity property for a complete
execution trace. We do this by requiring that trace integrity holds
for every subtrace delimited by a call and its corresponding return.

\definition Given a trace \(\MPT\) and a call map \(\callmap\)
%and a function map \(\codemap\),
an \(n\)-{\em transition} is a pair of
machine-policy state pairs \(\mpstate{\mach_c}{\pol_c}\) and
\(\mpstate{\mach_e}{\pol_e}\) in the trace such that
\(\mach_c\) is a valid call with \(n\) arguments, and
\(\mpstate{\mach_c}{\pol_c}\) steps to \(\mpstate{\mach_e}{\pol_e}\).

\apt{or is in some function?? and why do we need either??}:
\sna{I think Andrew is right that we don't need to know the function
     here, or that the entry point is valid; that's part of WBCF.}
%and \(\mach_e\) is an entry point for some function \(f\):

  \[\begin{split}
    \mathit{Tr}_n ~ \MPT ~ \mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}
    \triangleq & \exists \obs, f.
    ~ \mpstate{\mach_c}{\pol_c} \stepstopol{\obs} \mpstate{\mach_e}{\pol_e} \\
    & \land \callmap(\mach_c(\PCname)) = n
      %\land \codemap(\mach_e(\PCname)) = f
  \end{split}\]

\definition

{\em Stack integrity} is a property of a system with regard to a call map
\(\callmap\) that holds when every \(n\)-transition \(\mpstate{\mach_c}{\pol_c}\)
and \(\mpstate{\mach_e}{\pol_e}\) in every legal trace exhibits trace integrity
on the prefix up to its return, if any. For any initial state \(\mach_0\)
we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow
\MPT\). For each \(n\)-transition \(\mathit{Tr}_n ~ \MPT ~
\mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}\) we take the prefix of
the call up to its return, \(\mpstate{\mach_e}{\pol_e} \hookrightarrow
\MPT_{call} | \ret{\mach_c}\). If for any such call trace \(\MPT_{call}\), 
\(\propdef{TraceIntegrity} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
system as a whole enjoys stack integrity with regard to \(\callmap\).

\apt{Some more words needed here to smooth the technicalities.}
\sna{Attempt...}

@@

Unlike integrity, confidentiality is a noninterference-style property,
comparing the behavior of a machine with an arbitrary variation that
preserves low confidentiality data.

\definition%{Variations}
We say two machine states \(\mach\) and \(\mach'\) are {\em variations}
of one another given a contour \(\contour\) if they agree on low
confidentiality data:
%
\[
\mach \approx_\contour \mach' \triangleq \forall \component .
\contour(\component) = \LOWSEC \Rightarrow \mach(\component) =
\mach'(\component)
\]
%
Then, the trace version of confidentiality operates on two traces: the
``original'' trace $\MPT \in \MPOTRACE$ that corresponds to a callee $f$
and a ``variant'' trace $\machT \in \MTRACE$ that corresponds to how $f$
would behaves if it were started from a variant initial machine state.
%
As an example of variations, consider the top of \Cref{fig:confex1}
which shows a snapshot of the state as in the running example (top
memory layout) and a variation of it (bottom). \leo{Finalize
  contents, describe}

\definition

{\em Trace confidentiality} will be a property of two traces, \(\MPT\) and
\(\machT\), with regard to a return predicate \(R \subset \MACHS\) that
reflects the states that are considered to have returned. We abstract over
\(R\) because what it means to return depends on the state from which
\(f\) was called. There are three cases in which \(\MPT\) and \(\machT\)
enjoy trace confidentiality with respect to \(R\).

\sna{Interestingly, split out like this, we don't need to talk about
     contour here -- they are varied in the next property.}
\begin{itemize}
  \item If \(\MPT\) has a final state \(\last(\MPT) = \mpostate{\mach_r}
    {\pol_r}{\obs_r}\) and \(R\ \mach_r\), then \(\machT\) is also finite with
    some final state \(\mach_r'\) and any changes from the initial states are
    matched between \(\mach_r\) and \(\mach_r'\). In addition their
    observations are similar.
    \[\begin{split}
      \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \triangleq &
      \forall ~ \mach_r ~ \pol_r ~ \obs_r ~ \component . \last(\MPT) =
        \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
      & \exists \mach_r' ~ \obs_r' . \last(\machT) = (\mach_r',\obs_r')
        \land \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
      & ~ (\mach_e(\component) \not = \mach_r(\component)
        \lor \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
      \mach_r(\component) = \mach_r'(\component))
    \end{split}\]
  \item If \(\MPT\) is infinite, then \(\machT\) is also infinite and their
  observations are similar.
    \[\begin{split}
      \mathit{InfCase} ~ R ~ \MPT ~ \machT \triangleq
      & \neg (\exists mpo . \last(\MPT) = mpo) \Rightarrow \\
      & \neg (\exists mo . \last(\machT) = mo)
        \land \pi_o(\MPT) \eqsim \pi_o(\machT)
    \end{split}\]
\item If \(\MPT\) ends prematurely without fulfilling \(R\), then
  its observations are a prefix of those of \(\machT\).
    \[\begin{split}
      \mathit{FailCase} ~ R ~ \MPT ~ \machT \triangleq
      & \forall \mach_r ~ \pol_r ~ \obs_r .
        \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
      & \neg ~ R ~ \mach_r \Rightarrow \\
      & \pi_o(\MPT) \lesssim \pi_o(\machT)
    \end{split}\]
\end{itemize}

Putting it all together:

\[\begin{split}
  \propdef{TraceConfidentiality} ~ R ~ \MPT ~ \machT \triangleq
  & \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \land \\
  & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land
    \mathit{FailCase} ~ R ~ \MPT ~ \machT \\
\end{split}\]

In the non-return cases, we only care that the traces exhibit the same
observable behavior (or a prefix if the trace halts prematurely), as changes
in the state will not affect the caller absent a return.  In the return case
we must additionally guarantee that no changes were made based on secret data
that could corrupt the caller upon return.

Consider the result of modifying line 42 of our running example to $r_4 \leftarrow [\rsp - 9]$.
Instead of reading the proper argument {\tt v}, {\tt h} now reads directly from the
secret value of {\tt x} that belongs to {\tt main}. \Cref{fig:confex}
shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
in {\high} and low confidentiality in {\low}. Below it, the variant state
\(\mach_e'\) agrees on low confidentiality memory, but all other addresses
hold -5, in {\varied}.  After the last step, the return states \(\mach_r\) and
\(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
\(\mach_e'\). Addresses 108 and 109 have changed, shown in light {\low}, and while 109
had changed in a consistent way (independent of the variation), 108 holds either 43 or -4, so the trace does not obey trace confidentiality.
\rb{Color name macros}
\apt{Explain slash notation in effects column.}

As another example, 
changing line 30 of the running example to $r_5  \leftarrow [\rsp - 6]$ would cause
{\tt g} to read the value of {\tt main}'s local variable {\tt x} and then immediately
output it, violating the internal observation clause of the confidentiality definition.

@@

@@

\definition%{Stack Confidentiality}
%Let $\callmap$ be a call map and $\codemap$ a function map.
We will say that a system satisfies {\em stack confidentiality} with respect to
\(\callmap\) in a similar way to stack integrity. For every initial state
\(\mach_0\), we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)}
\hookrightarrow \MPT\). Then for any \(n\)-transition
\(\mathit{Tr}_n ~ \MPT ~ \mpstate{\mach_c}{\pol_c} ~
\mpstate{\mach_e}{\pol_e}\), we take the prefix trace until return:
%
\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{call} | \ret{\mach_c}\]

We also take the corresponding trace from any variant \(\mach_e' 
\approx_{\mathit{Cof}(\mach_c,n)} \mach_e\):
%
\[\mach_e' \hookrightarrow \machT_{call} | \ret{\mach_c}\]

If, for all such pairs of traces, \(\propdef{TraceConfidentialiy} ~
\ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
stack confidentiality with regard to \(\callmap\).

\apt{oy. still needs work to make this intelligible}
\sna{Attempted. There's a tension between introducing all the variables
     and keeping a good flow that I'm not sure I'm managing here though.}

@@

\definition
  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{stack integrity test} if their high-integrity components
  match:
  \[
    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\component) = \mach'(\component)
  \]

\definition
  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
    confidentiality test} if their observations and changed components coincide:
  \[
    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
  \]
\apt{scoping of $\forall$ is ambiguous.}
\rb{Fix notations, including for properties below.
  In particular, note that the last four arguments could be computed from the
  first two and it may be clearer to elide them.}
\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}

\definition The testing property keeps track of relevant information
on a {\em variant stack}: a list of machine states, each element of
which (denoted $\vm{C}$) of which is a variation of the actual machine
according to a contour $C$.
%
\footnote{The full formal definition in Coq also records additional
  information (the initial machine state at the call, its initial
  variation, the contour, and a return predicate) which we elide
  here for simplicity of exposition.}
%%
%  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
%  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
%  $$\vs \in \VSS$$
%%
\iflater
\leo{This can be elided}
  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
  respect to it if the stack is nonempty, the return map of the bottom element
  is the false property, and for each element $\vse \in \vs$:
  \begin{itemize}
  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
  \item $\mach$ is in the trace $\machT$ of its initial machine,
    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
  \item Its variant state $\mathit{curr\_variant}(\vse)$
    is in the trace of its initial variant,
    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item The registers of its contour $\contour(\vse)$ are public.
  \end{itemize}
  \fi

\definition
%
  Our testing property extends the basic tests above to whole variant
  stacks.  Given a machine that steps $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$, say that a variant stack $\vs$ \iflater that
  is well-formed with respect to $\mach$\fi safisfies the
  \emph{variant stack test} if for each variant element $\vm{C}$
  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step
  $\mathit{curr\_variant}(\vse) \stepstoobs{\obs} \mach'$ while
  satisfying the stack integrity test $\mathit{int}_T ~ \contour(\vse)
  ~ \mach ~ \mach'$ and stack confidentiality test $\mathit{conf}_T ~
  \mach ~ \mach'$ with regard to the original machine state. We write
  this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.

\definition{Eager testing property}
  Let $\callmap$ a call map and $\codemap$ a function map. Given a state
  $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
  say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
  testing property} at a given call depth if:

@@

  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow
    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
    $\mathit{stack}_T ~ d ~ [(\mach, \pol)] ~ \vs$, for any $d$ up to the size of
    $\vs$. (That is, a violation was detected $d$ levels into the stack.
    Elements are created and pushed to the variant stack on calls and popped
    from it on returns, so its size always reflects the depth of the program
    stack.)

  \item If the top variant of the stack (and only the top variant) returns,
    i.e., $\mathit{ret}(\head(\vs), \mach)$ holds, the property is safisfied for
    the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
    stack: $\mathit{stack}_T ~ |\vs| ~ [(\mach, \pol)] ~ \vs$. \rb{Explain the
    use of this vs the return case below.}
